Follow-up Samples

### 2017-02-07 submission/manuscript_outline_20161024.Rmd

***Full and Reduced Model***
Since differences were observed between initial and follow up samples and only a small number of individuals were positive for previously associated CRC bacteria; we next investigated if we could create models that could adequately classify and adjust either lesion or before sample probability based on the bacterial community and FIT.  The lesion model had an AUC range of `r format(min(lesion_AUC_run_summary$ROC), digits = 3)` to `r format(max(lesion_AUC_run_summary$ROC), digits = 3)` versus the before sample model which had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` to `r format(max(IF_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  Interestingly, identification of the most important variables and reducing the models to only these factors increased the AUC in the lesion model (`r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) and before sample model (`r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)`).

The test set AUC range for the full and reduced lesion model were similar to that reported for the training set AUC ranges and the ROC curve ranges overlap each other [Figure 3a].  The ROC curve for the final lesion model used was within the range of both the full and reduced lesion model [Figure 3a].  Interestingly, the test set AUC range for the before sample model performed much better then the training set AUCs.  Both the full and reduced before sample models overlapped with each other [Figure 3b] there was a marked decrease in the ROC curve for the final model used.

***Most Important Variables***
The reduced models were built based on the most important variables to the respective full model.  For the lesion model there were a total of `r length(lesion_top10_varsMDA$variable)` variables [Figure S2] whereas for the before sample model there were a total of `r length(IF_top10_varsMDA$variable)` variables [Figure S3].  For both models FIT resulted in the largest decrease in MDA [Figure S2a & S3a].

@@

***Common OTUs to both Models***
There were a total of `r length(common_vars_summary$otu)` OTUs that were common to both models.  Of these OTUs the most common taxonomic identifications were to  `r common_vars_summary['Otu000012', 'lowest_ID']`, `r common_vars_summary['Otu000005', 'lowest_ID']`, `r common_vars_summary['Otu000082', 'lowest_ID']`, and `r paste(gsub("_unclassified", "", common_vars_summary['Otu000006', 'lowest_ID']))`.  The majority of these OTUs had classifications to bacteria typically thought of as commensal [Table S2].

>>>>

***Full and Reduced Model***
We next wanted to identify if there were any common bacterial microbiome changes in individuals with adenoma or carcinoma.  In order to investigate this we created two different models: one to classify lesion versus normal and one to classify pre versus post samples based on the bacterial community and FIT.  The lesion model had an AUC range of `r format(min(lesion_AUC_run_summary$ROC), digits = 3)` to `r format(max(lesion_AUC_run_summary$ROC), digits = 3)` while the initial sample model had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` to `r format(max(IF_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  By identifying the most important variables for each respective model and then reducing them to only these factors we were able to increase the AUC in the lesion model (`r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) and initial sample model (`r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)`).        
The test set AUC range for the full and reduced lesion model were similar to that reported for the training set AUC ranges and the ROC curve ranges overlap each other [Figure 3a].  The ROC curve for the final lesion model used was within the range of both the full and reduced lesion model [Figure 3a].  Interestingly, the test set AUC range for the initial sample model performed much better then the training set AUCs.  Both the full and reduced initial sample models overlapped with each other [Figure 3b] but there was a marked decrease in the ROC curve for the final model used.

***Common OTUs to both Models***
The reduced models were built based on the most important variables to either classification model.  For the lesion model there were a total of `r length(lesion_top10_varsMDA$variable)` variables [Figure S2] whereas for the before sample model there were a total of `r length(IF_top10_varsMDA$variable)` variables [Figure S3].  For both models FIT resulted in the largest decrease in MDA [Figure S2a & S3a].  There were a total of `r length(common_vars_summary$otu)` OTUs that were common to both models.  Of these OTUs the most common taxonomic identifications were to  `r common_vars_summary['Otu000012', 'lowest_ID']`, `r common_vars_summary['Otu000005', 'lowest_ID']`, `r common_vars_summary['Otu000082', 'lowest_ID']`, and `r paste(gsub("_unclassified", "", common_vars_summary['Otu000006', 'lowest_ID']))`.  The majority of these OTUs had classifications to bacteria typically thought of as commensal [Table S2].

@@
